\MakeUppercase{Graph Product Structure for Non-Minor-Closed Classes}

### 2022-03-07 k-planar.tex
updates
More PDFsssssss

  The proof is by induction on the number of inner vertices of $N$, i.e., on $|V(N)\setminus V(F)|$.  If $N$ has no inner vertices, then we can simply take $\mathcal{P}:=\{P_1,\ldots,P_k\}$ and verify the preconditions of the lemma (the setup) ensure that $\mathcal{P}$ satisfies the requirements of the lemma.  Therefore, we may assume that $N$ has at least one inner vertex.

	First note that every inner face of $N'$ is bounded by three edges of $G'$. 
  Assign each vertex $v$ of $N'$ a colour $\alpha(v)\in\{1,2,3\}$ as follows: 
  Let $w$ be the first vertex of $F$ encountered on the path in $T$ from $v$ to the root of $T$.  (This implies that $w=v$ for $v\in V(F)$.)  Since $w\in V(F)$, $w\in P_i$ for some $i\in\{1,2,3\}$ and we define $\alpha(v):=i$.

  If $k<3$ then let $\tau:=v_1v_2v_3$ be any inner face of $N'$.  Otherwise, $k=3$, in which case we apply Sperner's Lemma implies to obtain an inner face $\tau:=v_1v_2v_3$ such that $\alpha(v_i)=i$ for each $i\in\{1,2,3\}$.  
  For each $i\in\{1,2,3\}$, let $Q_i$ be the shortest path, in $T$, from $v_i$ to $V(F)$.
   % 
   % 
   % If $k=3$, set $R_i := P_i$ for each $i\in\{1,2,3\}$.  Otherwise, as in \citep{DJMMUW20}, split $P_1,\ldots,P_k$ to partition $V(F)$ into three sets $R_1$, $R_2$, and $R_3$ such that each $F[R_i]$ is a non-empty path and each $R_i$ contains vertices from exactly one of $P_1,\ldots,P_k$.
  % 
	% Next, as in \citep{DJMMUW20}, use Sperner's Lemma to find an inner face $\tau=v_1v_2v_3$ of $N'$ such that, $T$ contains disjoint vertical paths $Q_1,Q_2,Q_3$ such that each $Q_i$ begins at $v_i$, ends at some vertex in $R_i$, and whose internal vertices (if any) are contained in $N'-V(F)$.
	Let $\overline{Y}$ denote the subgraph of $N'$ consisting of vertices and edges of $Q_1$, $Q_2$, $Q_3$, and $\tau$.  Let $\overline{Y}^+$ denote the subgraph of $N$ consisting of the vertices and edges of $\overline{Y}$ plus the vertices and edges of every kite formed by a crossing between an edge of $G$ and an edge of $\overline{Y}$.

@@

	Let $M$ and $M^+$ denote the subgraph of $G$ containing the edges and vertices of $\overline{Y}$, respectively $\overline{Y}^+$, and the edges and vertices of $F$.  The graph $M^+$ has some number of bounded faces, all contained in the interior of $F$. Some of the bounded faces of $M^+$ are kite faces. Let $F_1,\ldots,F_m$ be the non-kite bounded faces of $M^+$.

	We claim that, for each $i\in\{1,\ldots,m\}$, the boundary of $F_i$ is a cycle in $G'$ that contains no spars. Otherwise, some edge $vw$ contributes to the boundary of $F_i$ but is crossed by an edge $xy\in E(G)$. Then, $vw\not\in E(F)$ since no edge of $F$ is crossed by any edge of $G$. Therefore $vw\in E(\overline{Y}^+)$ so $xy\in E(Y^+)$. But then the only faces of $M^+$ incident to $vw$ are kite faces.  In particular $vw$ cannot be incident to the non-kite face $F_i$.

	Observe that each of the faces $F_1,\ldots,F_m$ is contained in a single internal face of $M$.   Let $Y^+ := \overline{Y}^+-F$. Therefore, $V(F_i)$ can be partitioned into at most three sets $P_1'$, $P_2'$, and $P_3'$ where $P_1'\subset V(Y^+)$, $P_2'\subseteq P_a$, $P_3'\subseteq P_b$ for some $a,b\in\{1,2,3\}$, and $F_i[P_j']$ is a path, for each $j\in\{1,2,3\}$.

>>>>

  The proof is by induction on the number of inner vertices of $N$, i.e., on $|V(N)\setminus V(F)|$.  If $N$ has no inner vertices, then we can simply take $\mathcal{P}:=\{P_1,\ldots,P_k\}$ and verify the preconditions of the lemma (the `setup') ensure that $\mathcal{P}$ satisfies the requirements of the lemma.  Therefore, we may assume that $N$ has at least one inner vertex.

  Assign each vertex $v$ of $N'$ a colour $\alpha(v)\in\{1,2,3\}$ as follows: 
  Let $w$ be the first vertex of $F$ encountered on the path in $T$ from $v$ to the root of $T$.  (This implies that $w=v$ for $v\in V(F)$.)  Since $w\in V(F)$, $w\in P_i$ for some $i\in\{1,2,3\}$ and we define $\alpha(v):=i$.

  If $k<3$ then let $\tau:=v_1v_2v_3$ be any inner face of $N'$.  Otherwise, $k=3$, in which case we apply Sperner's Lemma to obtain an inner face $\tau:=v_1v_2v_3$ such that $\alpha(v_i)=i$ for each $i\in\{1,2,3\}$.  
  For each $i\in\{1,2,3\}$, let $Q_i$ be the shortest path, in $T$, from $v_i$ to $V(F)$.
   % 
   % 
   % If $k=3$, set $R_i := P_i$ for each $i\in\{1,2,3\}$.  Otherwise, as in \citep{DJMMUW20}, split $P_1,\ldots,P_k$ to partition $V(F)$ into three sets $R_1$, $R_2$, and $R_3$ such that each $F[R_i]$ is a non-empty path and each $R_i$ contains vertices from exactly one of $P_1,\ldots,P_k$.
  % 
	% Next, as in \citep{DJMMUW20}, use Sperner's Lemma to find an inner face $\tau=v_1v_2v_3$ of $N'$ such that, $T$ contains disjoint vertical paths $Q_1,Q_2,Q_3$ such that each $Q_i$ begins at $v_i$, ends at some vertex in $R_i$, and whose internal vertices (if any) are contained in $N'-V(F)$.
	Let $\overline{Y}$ denote the subgraph of $N'$ consisting of vertices and edges of $Q_1$, $Q_2$, $Q_3$, and $\tau$.  Let $\overline{Y}^+$ denote the subgraph of $N$ consisting of the vertices and edges of $\overline{Y}$ plus the vertices and edges of every kite formed by a crossing between an edge of $G$ and an edge of $\overline{Y}$.  Let $Y:= V(\overline{Y}-F)$ and let $Y^+:=V(\overline{Y}^+-F)$.

@@

	Let $M$ and $M^+$ denote the subgraph of $G$ containing the edges and vertices of $\overline{Y}$, respectively $\overline{Y}^+$, and the edges and vertices of $F$.  By the choice of $\tau$, the vertices on the boundary of each inner face of $M$ are contained in $Y\cup P_a\cup P_b$ for some $a,b\in\{1,\ldots,k\}$.  In particular if $k=3$ then, for each inner face $f$ of $M$ there exists some $c\in\{1,2,3\}$ such that no vertex of $P_c$ is on the boundary of $f$.

  The graph $M^+$ has some number of inner faces, all contained in the interior of $F$. Some of the inner faces of $M^+$ are Type~(ii) faces. Let $F_1,\ldots,F_m$ be the Type~(i) inner faces of $M^+$.

	We claim that, for each $i\in\{1,\ldots,m\}$, the boundary of $F_i$ is a cycle in $G'$ that contains no spars. Otherwise, some edge $vw$ contributes to the boundary of $F_i$ but is crossed by an edge $xy\in E(G)$. Then, $vw\not\in E(F)$ since no edge of $F$ is crossed by any edge of $G$. Therefore $vw\in E(\overline{Y}^+)$ so $xy\in E(Y^+)$. But then the only faces of $M^+$ incident to $vw$ are Type~(ii) faces.  In particular $vw$ cannot be incident to the Type~(i) face $F_i$.

	Observe that each of the faces $F_1,\ldots,F_m$ is contained in a single inner face of $M$.   Let $Y^+ := \overline{Y}^+-F$. Therefore, $V(F_i)$ can be partitioned into at most three sets $P_1'$, $P_2'$, and $P_3'$ where $P_1'\subset V(Y^+)$, $P_2'\subseteq P_a$, $P_3'\subseteq P_b$ for some $a,b\in\{1,2,3\}$, and $F_i[P_j']$ is a path, for each $j\in\{1,2,3\}$.
